\LaTeX\ Author Guidelines for ICCV Proceedings

### 2021-03-10 chapters/0_abstract.tex
Rewrite supplementary material and add proofs

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a new simple theoretical framework that generalizes algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines GASP, a Generalized Algorithm for Signed graph Partitioning, and allows us to explore many combinations of different linkage criteria and cannot-link constraints. 
We prove the equivalence of existing clustering methods to some of those combinations, and introduce new algorithms for combinations which have not been studied before. 
We also study theoretical properties of these combinations and prove which one of them define ultrametrics on the graph.
% On stochastic block model problems, GASP compares favorably to spectral clustering. 
We conduct a systematic comparison of various instantiations of GASP on a large variety of both synthetic and existing signed clustering problems, in terms of accuracy but also efficiency and robustness to noise. 
More importantly, we find that some of the algorithms proposed in our framework, when combined to the predictions from a CNN model, result in a simple instance segmentation pipeline that achieves state-of-the-art performances on the competitive CREMI 2016 EM segmentation benchmark.
% \keywords{Partitioning of Signed Graphs, Generalized Framework, Agglomerative Clustering, Instance Segmentation}
\end{abstract}

>>>>

% !TEX root = ../agglo_clust_review.tex
%%%%%%%%% ABSTRACT
\begin{abstract}
We propose a new simple theoretical framework that generalizes algorithms for hierarchical agglomerative clustering to weighted graphs with both attractive and repulsive interactions between the nodes. This framework defines GASP, a Generalized Algorithm for Signed graph Partitioning, and allows us to explore many combinations of different linkage criteria and cannot-link constraints. 
We prove the equivalence of existing clustering methods to some of those combinations and introduce new algorithms for combinations that have not been studied before. 
We study both theoretical and empirical properties of these combinations and prove which ones of them define an ultrametric on the graph.
We conduct a systematic comparison of various instantiations of GASP on a large variety of both synthetic and existing signed clustering problems, in terms of accuracy but also efficiency and robustness to noise. 
Lastly, we find that some of the algorithms included in our framework, when combined with the predictions from a CNN model, result in a simple instance segmentation pipeline that achieves state-of-the-art performances on the competitive CREMI 2016 EM segmentation benchmark.

% On stochastic block model problems, GASP compares favorably to spectral clustering. 
% \keywords{Partitioning of Signed Graphs, Generalized Framework, Agglomerative Clustering, Instance Segmentation}
\end{abstract}

### 2021-03-10 chapters/1_intro.tex
Rewrite supplementary material and add proofs

\begin{figure*}[t]
\centering
\includegraphics[width=0.98\textwidth]{figs/intro_image_v5.pdf} % left bottom right top
\caption{\textbf{(a)} Some iterations of \algname{}, the proposed Generalized Agglomerative Algorithm for Signed Graph Partitioning, on a toy graph with attractive/positive (green) and repulsive/negative (red) interactions. At each iteration, the yellow edge with highest interaction is contracted (average and sum linkage methods are shown). \textbf{(b)} List of linkage criteria used in this paper (see definitions in Table~\ref{tab:linkage-criteria} below), demonstrated on two small clusters.  \textbf{(c)} Application of \algname{} to instance segmentation: we show raw data from the CREMI 2016 neuron-segmentation challenge and some predictions of our CNN model, where white pixels represent boundary evidence. \textbf{(d)} \algname{} agglomeration order for sum and average linkage: we show which pairs of neighboring pixels were merged first (white), later on (brown/red), or never (black).
\label{fig:intro_figure}}
\end{figure*}

@@

            \thead[r]{GASP on positive-weighted graphs} & \thead{-} &\thead{\textbf{HC-Single}} &\thead{\textbf{HC-Avg}} &\thead{\textbf{HC-Single}} &\thead{\textbf{HC-Complete}} \\
            \thead[r]{GASP on signed graphs} & \thead{GAEC \cite{keuper2015efficient}} & \thead{\textbf{Mutex Watershed} \cite{wolf2018mutex}}& \thead{\textbf{HC-Avg}} &\thead{\textbf{HC-Single}} &\thead{\textbf{HC-Complete}} \\
            \thead[r]{GASP on signed graphs + constraints} & \thead{\colorbox{yellow}{HCC-Sum}} % \thead{Greedy Fixation \cite{levinkov2017comparative}} 
            & \thead{\textbf{Mutex Watershed} \cite{wolf2018mutex}}& \thead{\colorbox{yellow}{HCC-Avg}} &  \thead{\colorbox{yellow}{HCC-Single}} &  \thead{\colorbox{yellow}{HCC-Complete}} \\
            % \multicolumn{2}{c|}{\multirow{2}{*}[-0.5em]{\thead{\textbf{\algname{} linkage criteria} $\,\,\interact(S_u ,S_v)$}}}  & \multirow{2}{*}[-0.5em]{\thead{\textbf{Unsigned Graphs}}} & \multicolumn{2}{c}{\thead{\textbf{Signed Graphs}}}  \\        
            % \multicolumn{2}{c|}{} &  &  \multicolumn{1}{c}{\thead{No Constraints}} & \thead{With Constraints} \\ \midrule

@@

        \end{tabular}
    \end{subtable} 
    \caption{Clustering algorithms included in the proposed \algname{} framework, given a linkage criterion, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. New algorithms are highlighted in yellow (hierarchical clustering with constraints, HCC). Algorithms typeset in bold font define an ultrametric on the graph (see Eq.~\ref{eq:UM_def}). Algorithms included in the \TODO{green box}, are weight-shift invariant (see Prop.~\ref{prop:weight_shift_invariant}). 
    We denote as $E_{ij}=(S_i \times S_{j}) \cap E$ the set of edges connecting two clusters $S_i, S_j$. } 
    \label{tab:linkage-criteria}
\end{table*}

@@

In this paper, we propose a simple theoretical framework that generalizes over agglomerative algorithms for signed graphs by linking them to hierarchical clustering (HC) on positive-weighted graphs \cite{lance1967general}. This framework defines an underlying basic algorithm and allows us to explore its combinations with different linkage criteria and \emph{cannot-link constraints} (see Fig.~\ref{fig:intro_figure}a-b and Table~\ref{tab:linkage-criteria}). 
We then formally prove that some of the combinations correspond to existing clustering algorithms, and introduce new algorithms for combinations which have not been explored before. By analyzing their theoretical properties, we also show which one of them define ultrametrics on the graph.

We evaluate the algorithms on a large variety of existing and synthetic signed graph clustering problems. 
Moreover, we evaluate all algorithms on \emph{instance segmentation} -- a computer vision task consisting in assigning each pixel of an image to an object instance -- by predicting graph edge weights with a CNN (see Fig.~\ref{fig:intro_figure}).
Our experiments show that the choice of a linkage criterion strongly influences the way how clusters are grown by the agglomerative algorithms, making some linkage methods more suited for certain types of clustering problems.
% We focus our comparison on three of the best-performing algorithms included in our framework, which are based on average, sum, and absolute maximum linkage criteria (see Table~\ref{tab:linkage-criteria}). We highlight their empirical properties by showing how clusters are grown differently depending on the algorithm.
We benchmark the clustering algorithms by focusing on their efficiency, robustness and tendency to over- or under-cluster. 
First, we observe that the tested agglomerative algorithms strongly outperform recently proposed spectral clustering methods. Then,  we also show that average-linkage based agglomerative algorithms achieve state of the art results on the competitive CREMI 2016 challenge for neuron segmentation of 3D electron microscopy image volumes of brain tissue.

>>>>

\begin{figure*}[t]
\centering
\includegraphics[width=0.98\textwidth]{figs/intro_image_v5.pdf} % left bottom right top
\caption{\textbf{(a)} Some iterations of \algname{}, the proposed Generalized Agglomerative Algorithm for Signed Graph Partitioning, on a toy graph with attractive/positive (green) and repulsive/negative (red) interactions. At each iteration, the yellow edge with highest interaction is contracted. Average and sum linkage methods are shown. \textbf{(b)} List of linkage criteria used in this paper (see definitions in Table~\ref{tab:linkage-criteria} below), demonstrated on two small clusters.  \textbf{(c)} Application of \algname{} to instance segmentation: we show raw data from the CREMI 2016 neuron-segmentation challenge and some predictions of our CNN model, where white pixels represent boundary evidence. \textbf{(d)} \algname{} agglomeration order for sum and average linkage: we show which pairs of neighboring pixels were merged first (white), later on (brown/red), or never (black).
\label{fig:intro_figure}}
\end{figure*}

@@

            \thead[r]{GASP on positive-weighted graphs} & \thead{-} &\tikzmark{a} \thead{\textbf{HC-Single}} &\tikzmark{g} \thead{\textbf{HC-Avg}} &\thead{\textbf{HC-Single}}\tikzmark{z} &\thead{\textbf{HC-Complete}} \tikzmark{b} \\
            \thead[r]{GASP on signed graphs} & \thead{GAEC \cite{keuper2015efficient}} &  \thead{\textbf{Mutex Watershed} \cite{wolf2018mutex}}& \tikzmark{f} \thead{\textbf{HC-Avg}} &\thead{\textbf{HC-Single}} \tikzmark{e} &\thead{\textbf{HC-Complete}}\tikzmark{w} \\
            \thead[r]{GASP on signed graphs + constraints} & \thead{\colorbox{yellow}{HCC-Sum}} % \thead{Greedy Fixation \cite{levinkov2017comparative}} 
            & \thead{\textbf{Mutex Watershed} \cite{wolf2018mutex}}& \thead{\colorbox{yellow}{HCC-Avg}} &  \thead{\colorbox{yellow}{HCC-Single}} \tikzmark{d} &   \thead{\textbf{HC-Complete}} \tikzmark{c} \\
            % \multicolumn{2}{c|}{\multirow{2}{*}[-0.5em]{\thead{\textbf{\algname{} linkage criteria} $\,\,\interact(S_u ,S_v)$}}}  & \multirow{2}{*}[-0.5em]{\thead{\textbf{Unsigned Graphs}}} & \multicolumn{2}{c}{\thead{\textbf{Signed Graphs}}}  \\        
            % \multicolumn{2}{c|}{} &  &  \multicolumn{1}{c}{\thead{No Constraints}} & \thead{With Constraints} \\ \midrule

@@

        \end{tabular}
    \end{subtable} 
    \caption{Clustering algorithms included in the proposed \algname{} framework, given a linkage criterion, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. New constrained hierarchical clustering algorithms (HCC) are highlighted in yellow. Algorithms typeset in bold-font define an ultrametric on the graph (Eq.~\ref{eq:UM_def}). Algorithms included in the green box, are weight-shift invariant (Prop.~\ref{prop:weight_shift_invariant}). 
    Given a graph $\mathcal{G}(V,E)$, we denote as $E_{ij}=(S_i \times S_{j}) \cap E$ the set of edges connecting two clusters $S_i \subseteq V$, $S_j \subseteq V$. } 
    \label{tab:linkage-criteria}
\end{table*}

@@

In this paper, we propose a simple theoretical framework that generalizes over agglomerative algorithms for signed graphs by linking them to hierarchical clustering (HC) on positive-weighted graphs \cite{lance1967general}. This framework defines an underlying basic algorithm and allows us to explore its combinations with different linkage criteria and \emph{cannot-link constraints} (see Fig.~\hyperref[fig:intro_figure]{\ref*{fig:intro_figure}a}, \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}b}, and Table~\ref{tab:linkage-criteria}). 
We then formally prove that some of the combinations correspond to existing clustering algorithms, and introduce new algorithms for combinations which have not been explored before. By analyzing their theoretical properties, we also show which one of them define an ultrametric on the graph.

We evaluate the algorithms on a large variety of both existing and synthetically generated signed graph clustering problems. 
Next, we also test the algorithms on \emph{instance segmentation} -- a computer vision task consisting in assigning each pixel of an image to an object instance -- by predicting graph edge weights with a CNN (Fig.~\hyperref[fig:intro_figure]{\ref*{fig:intro_figure}c}).
Our experiments show that the choice of a linkage criterion strongly influences the way how clusters are grown by the agglomerative algorithms (Fig.~\hyperref[fig:intro_figure]{\ref*{fig:intro_figure}d}), making some linkage methods more suited for certain types of clustering problems.
% We focus our comparison on three of the best-performing algorithms included in our framework, which are based on average, sum, and absolute maximum linkage criteria (see Table~\ref{tab:linkage-criteria}). We highlight their empirical properties by showing how clusters are grown differently depending on the algorithm.
We benchmark the clustering algorithms by focusing on their efficiency, robustness and tendency to over- or under-cluster. 
On instance segmentation, we show that the tested agglomerative algorithms strongly outperform recently proposed spectral clustering methods, and that average-linkage based agglomerative algorithms achieve state of the art results on the competitive CREMI 2016 challenge for neuron segmentation of 3D electron microscopy image volumes of brain tissue.

\begin{tikzpicture}[remember picture,overlay]
% (shift_along_horizontal, shift_along_vertical)
% a ---------------------------- b
%            g -----------------
%            f         e ------- 
%                      d         c
% 
\draw [green,line]($(a)+(0,2ex)$)--($(b)+(0.9ex,2ex)$)--($(c)+(0.9ex,-1.5ex)$) -- ($(d)+(1.4ex,-1.5ex)$) -- ($(e)+(1.4ex,-1.2ex)$) -- ($(f)+(-0.3ex,-1.2ex)$)--($(g)+(-0.3ex,-1.2ex)$)--($(a)+(0,-1.2ex)$)--cycle;
\end{tikzpicture}

### 2021-03-10 chapters/3_algorithm.tex
Rewrite supplementary material and add proofs

\paragraph{Sum Linkage} -- 
An agglomerative algorithm with sum linkage does only make sense for graphs with signed edge weights, because only for these graphs the sum of two attractive (or repulsive) interactions is still attractive (repulsive). On positive weighted graphs, on the other hand, a strong attractive interaction could be obtained by summing many weak interactions, which is not desirable. Such an algorithm was pioneered in \cite{levinkov2017comparative,keuper2015efficient} and was named Greedy Agglomerative Edge Contraction (GAEC)\footnote{An algorithm equivalent to GAEC was recently independently re-proposed in \cite{chehreghani2020hierarchical}.}. All variations of \algname{} decrease the multicut objective defined in Eq.~\ref{eq:MC_objective} each time two clusters with positive interaction are merged. But only GAEC always makes the \emph{greedy-choice} that will mostly decrease the multicut objective at every step. The authors of \cite{levinkov2017comparative} also proposed a version of GAEC with cannot-link constraints named \emph{GreedyFixation}, but the constrained sum-link version of \algname{} proposed here (named \emph{HCC-Sum} in Table~\ref{tab:linkage-criteria}) did perform better in our experiments.

\paragraph{AbsMax Linkage} -- This linkage method is also specific to signed graphs, since on positive-weighted graphs it would be equivalent to single linkage. Here, we prove that the Mutex Watershed Algorithm \cite{wolf2018mutex} can be seen as an agglomerative algorithm with AbsMax linkage (see proof in Appendix):
\begin{prop} 
The \algname{} Algorithm \ref{main_alg} with AbsMax linkage (see def. in Table \ref{tab:linkage-criteria}) returns the same final clustering $\Pi^*_{\mathrm{AbsMax}}$ whether or not cannot-link constraints are introduced. In particular, $\Pi^*_{\mathrm{AbsMax}}$ is equal to the clustering returned by the Mutex Watershed Algorithm (MWS) \cite{wolf2018mutex}, which has empirical complexity $\mathcal{O}(N \log N)$. 
\end{prop}

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been throughly studied on positive-weighted graphs, but never on signed graphs. In the recent related work \cite{chehreghani2020hierarchical}, the authors prove that applying these three linkage methods to a signed graph is equivalent to applying them to the positive-weighted graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms included in our generalized framework have this invariance-property: (proof in Appendix)
\begin{prop} \label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the structure of the merging tree $T$ returned by such algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the versions of \algname{} studied in this paper, only those using Average, Single, and Complete linkage criteria without cannotLinkConstraints are weight-shift-invariant (see algorithms typeset in bold font in Table~\ref{tab:linkage-criteria}).
\end{prop}
\noindent To our knowledge, Avg, Single, and Complete linkage methods have never been combined with cannot-link constraints on signed graphs, so the algorithms that in Table~\ref{tab:linkage-criteria} are named \emph{HCC-Avg}, \emph{HCC-Single}, and \emph{HCC-Complete} are new.

@@

\paragraph{Algorithms defining an ultrametric} -- The connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known \cite{johnson1967hierarchical,milligan1979ultrametric}. Usually, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. Here, we generalize this concept to signed graphs. To do so, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``pseudo-distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. We then prove the following two propositions (proofs are left in Appendix):
\begin{prop}
Consider a graph $\mathcal{G}(V,E,w_e)$, a linkage criterion $\interact{}$, and an AC algorithm returning a binary rooted tree $T$ with height $h_T$. The quantity $(V, d_{T})$ defined in Eq.~\ref{eq:UM_def} is an ultrametric iff, $\forall u,v,t \in V$:
\begin{equation}
\interact{}_{T}(u,v) > \interact{}_{T}(u,t) \Leftrightarrow h_T(u \vee v) < h_T[u \vee t] 
\end{equation}
In words, $(V, d_{T})$ is an ultrametric iff the interaction $\interact{}_{T}(u, v)$ associated to the least common ancestor ${(u\vee v)\in T}$ is both lower than any other $(u\vee v)$-descendant node and higher than any other $(u\vee v)$-ancestor node in the tree $T$.
\end{prop}
\begin{prop}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms listed in Table \ref{tab:linkage-criteria} and included in the \algname{} framework, only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric given by the quantity $(V, d_{T^*})$ defined in Eq.~\ref{eq:UM_def}, where $T^*$ is the tree returned by Algorithm~\ref{main_alg}.
\end{prop}

>>>>

\paragraph{Sum Linkage} -- 
On signed graphs, the sum of two attractive (or repulsive) interactions is still attractive (repulsive). On the other hand, on positive weighted graphs, a strong attractive interaction could be obtained by summing many weak interactions, which is often not desirable. This explains why, to our knowledge, an agglomerative algorithm with sum linkage has never been used on positive weighted graphs. On signed graphs, such an algorithm was pioneered in \cite{levinkov2017comparative,keuper2015efficient} and was named Greedy Agglomerative Edge Contraction (GAEC)\footnote{An algorithm equivalent to GAEC was recently independently re-proposed in \cite{chehreghani2020hierarchical}.}. All variations of \algname{} decrease the multicut objective defined in Eq.~\ref{eq:MC_objective} each time two clusters with positive interaction are merged. But only GAEC always makes the \emph{greedy-choice} that will mostly decrease the multicut objective at every step. The authors of \cite{levinkov2017comparative} also proposed a version of GAEC with cannot-link constraints named \emph{GreedyFixation}, but the constrained sum-link version of \algname{} proposed here (named \emph{HCC-Sum} in Table~\ref{tab:linkage-criteria}) did perform better in our experiments.

\paragraph{AbsMax Linkage} -- This linkage method is also specific to signed graphs, since on positive-weighted graphs it would be equivalent to single linkage. Here, we prove that the Mutex Watershed Algorithm \cite{wolf2018mutex} can be seen as an agglomerative algorithm with AbsMax linkage (proofs of the following propositions are given in Appendix \ref{sec:proposition_proofs}):
\begin{restatable}{prop}{absmaxmutex}
\label{prop:absmax_mutex}
The \algname{} Algorithm \ref{main_alg} with AbsMax linkage (see def. in Table \ref{tab:linkage-criteria}) returns the same final clustering $\Pi^*_{\mathrm{AbsMax}}$ whether or not cannot-link constraints are introduced. In particular, $\Pi^*_{\mathrm{AbsMax}}$ is equal to the clustering returned by the Mutex Watershed Algorithm (MWS) \cite{wolf2018mutex}, which has empirical complexity $\mathcal{O}(N \log N)$. 
\end{restatable}
% \begin{prop} 
% \end{prop}

\paragraph{Average, Single, and Complete Linkage} -- These three linkage criteria have been throughly studied on positive-weighted graphs, but never on signed graphs. In the recent related work \cite{chehreghani2020hierarchical}, the authors prove that applying these three linkage methods to a signed graph is equivalent to applying them to the positive-weighted graph obtained by shifting all edge weights by a constant.  Here, we prove which of the algorithms included in our generalized framework have this invariance-property:
\begin{restatable}{prop}{invariantAlgs}
\label{prop:weight_shift_invariant}
We call an agglomerative algorithm ``weight-shift invariant'' if the structure of the merging tree $T$ returned by such algorithm is invariant w.r.t. a shift of all edge weights $w_e$  by a constant $\alpha\in \mathbb{R}$. Among the versions of \algname{} studied in this paper, only hierarchical clustering using Average (HC-Avg), Single (HC-Single), and Complete linkage (HC-Complete) are weight-shift-invariant (see algorithms included in the green box in Table~\ref{tab:linkage-criteria}).
\end{restatable}
\noindent To our knowledge, Average and Single linkage methods have never been combined with cannot-link constraints on signed graphs, so the constrained hierarchical clustering algorithms that in Table~\ref{tab:linkage-criteria} are named \emph{HCC-Avg} and \emph{HCC-Single} are new. In Lemma \ref{lemma:absMax_and_complete_property} (Appendix \ref{sec:proposition_proofs}), we prove that with a Complete linkage the agglomeration order does not change whether constraints are introduced or not.

@@

\paragraph{Algorithms defining an ultrametric} -- The connection between agglomerative algorithms and ultrametrics\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.} is well known \cite{johnson1967hierarchical,milligan1979ultrametric}. Usually, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. Here, we generalize this concept to signed graphs. To do so, we first map the signed interaction $\treeHeight{}$ defined in Eq.~\ref{eq:def_dendr_interact} to positive ``pseudo-distances'' $d_{T}:V \times V \rightarrow \mathbb{R}^{+}$:
\begin{align}\label{eq:UM_def}
d_{T}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv  \epsilon + \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v')
\end{align}
and where $\epsilon > 0$. By definition, we have that $d_{T}(u,v)=d_{T}(v,u)$ and $d_{T}(u,u)= 0$. We then prove the following two propositions:
\begin{restatable}{prop}{firstUltraMetricProperty}
\label{prop:ultraMetric1}
Consider a graph $\mathcal{G}(V,E,w_e)$, a linkage criterion $\interact{}$, and an agglomerative algorithm returning a binary rooted tree $T$ with height $h_T$. Then, $(V, d_{T})$ defined in Eq.~\ref{eq:UM_def} is an ultrametric if and only if the following is true
 % iff the interaction $\interact{}_{T}(u, v)$ associated to the least common ancestor ${(u\vee v)\in T}$ is both lower than the interaction of any other $(u\vee v)$-descendant node and higher than the interaction of any other $(u\vee v)$-ancestor node in the tree $T$. In formulae, this is equivalent to saying that, $\forall u,v,t \in V$:
\begin{equation}\label{eq:UM_assumption}
 h_T(u \vee v) < h_T(u \vee t) \Rightarrow \interact{}_{T}(u,v) \geq \interact{}_{T}(u,t),
 % h_T(u \vee v) < h_T(u \vee t) \Leftrightarrow \interact{}_{T}(u,v) > \interact{}_{T}(u,t). 
\end{equation}
which in words means: if the algorithm merges nodes $u,v$ before to merge nodes $u,t$, then the signed interaction $\interact{}_{T}(u,v)$ between $u$ and $v$ has to be higher or equal than $\interact{}_{T}(u,t)$.
% In words, $(V, d_{T})$ is an ultrametric iff the interaction $\interact{}_{T}(u, v)$ associated to the least common ancestor ${(u\vee v)\in T}$ is both lower than any other $(u\vee v)$-descendant node and higher than any other $(u\vee v)$-ancestor node in the tree $T$.
\end{restatable}
\begin{restatable}{prop}{secondUltraMetricProperty}
\label{prop:ultraMetric2}
% Consider the binary rooted tree $T^*$ returned by the \TODO{all variants in Table 1} \algname{} Algorithm~\ref{main_alg}, given a choice of linkage criteria $\interact{}$ and the use (or not) of CannotLinkConstraints. 
Among the algorithms included in the \algname{} framework (see Table \ref{tab:linkage-criteria}), only Mutex Watershed and hierarchical clustering with Average (HC-Avg), Single (HC-Single) and Complete linkage (HC-Complete) define an ultrametric $(V, d_{T^*})$, where $d_{T^*}$ is defined in Eq.~\ref{eq:UM_def} and $T^*$ is the tree returned by the \algname{} Algorithm~\ref{main_alg}.
\end{restatable}
